%\documentclass[12pt,draftcls]{ucdavisthesis}
\documentclass[12pt]{ucdavisthesis}

% PLEASE READ THE MANUAL - ucdavisthesis.pdf (in the package installation directory)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%               LATEX COMMANDS FOR DOCUMENT SETUP                      %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\usepackage{bookmark}
\usepackage[us,nodayofweek,12hr]{datetime}
\usepackage{graphicx}
\usepackage{color}
\usepackage{lineno}
\linenumbers*[1]
\usepackage{url}
\usepackage{amsmath}
\usepackage{rotating} 
\usepackage[bottom]{footmisc}
\usepackage{graphics}

\usepackage{color}
\definecolor{pinkred}{rgb}{1.0, 0.4, 0.4}
\usepackage[table]{xcolor}
%\usepackage[square,comma,numbers,sort&compress]{natbib}
%\usepackage{hypernat}
% Other useful packages to try
%\usepackage{amsmath}
%\usepackage{amssymb}
%
% Different fonts to try (uncomment only fontenc and one font at a time)
% (you may need to install these first)
%\usepackage[T1]{fontenc} %enable fontenc package if using one of the fonts below
%\usepackage[adobe-utopia]{mathdesign}
%\usepackage{tgschola}
%\usepackage{tgbonum}
%\usepackage{tgpagella}
%\usepackage{tgtermes}
%\usepackage{fourier}
%\usepackage{fouriernc}
%\usepackage{kmath,kerkis}
%\usepackage{kpfonts}
%\usepackage[urw-garamond]{mathdesign}
%\usepackage[bitstream-charter]{mathdesign}
%\usepackage[sc]{mathpazo}
%\usepackage{mathptmx}
%\usepackage[varg]{txfonts}

\hyphenation{dis-ser-ta-tion blue-print man-u-script pre-par-ing} %add hyphenation rules for words TeX doesn't know


%\renewcommand{\rightmark}{\scriptsize A University of California Davis\ldots \hfill Rev.~\#1.0 \quad Compiled: \currenttime, \today}
% a fancier running header that can be used with draftcls options

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%        DOCUMENT SETUP AND INFORMATION FOR PRELIMINARY PAGES          %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title          {Studying Regional Climate with Variable-Resolution CESM}
%Exact title of your thesis. Indicate italics where necessary by underlining or using italics. Please capitalize the first letter of each word that would normally be capitalized in a title.

\author         {Xingying Huang}
%Your full name as it appears on University records. Do not use initials.

\authordegrees  {B.S. (Wuhan University) 2010 \\
                 M.S. (Beijing Normal University) 2013}
%Indicate your previous degrees conferred.

\officialmajor  {Atmospheric Science}
%This is your official major as it appears on your University records.

\graduateprogram{Atmospheric Science}
%This is your official graduate program name. Used for UMI abstract.

\degreeyear     {2016}
% Indicate the year in which your degree will be officially conferred.

\degreemonth    {December}
% Indicate the month in which your degree will be officially conferred. Used for UMI abstract.

\committee{Professor Paul Ullrich, Chair}{Professor Terry Nathan}{Professor Daniel Feldman}{}{}
% These are your committee members. The command accepts up to five committee members so be sure to have five sets of braces, even if there are empties.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\copyrightyear{2020}
%\nocopyright

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%`%%%%

\dedication{\textsl{To someone very important \ldots \\
            a nice dedication.}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\abstract{Regional climate is becoming one of the most important areas under the umbrella of climate change research, as more information is now needed at finer scales.  High horizontal resolution is needed to allow a more accurate representation of fine scale forcing, and leads to a better representation of processes and interactions that are significantly drivers for regional and local climate variability. Regional climate models (RCMs) are a traditional method for modeling regional climate, and have been proven to be capable of both mean climate and extreme events studies. Over the past decade, variable-resolution global climate models (VRGCMs) have been introduced as an alternative way for studying regional climate and applications. In this research, the newly developed variable-resolution technique within the Community Earth System Model (CESM) has been applied for long-term regional climate studies. 

In this thesis, firstly, VR-CESM is assessed for long-term regional climate modeling over California against a traditional RCM -- the Weather Research and Forcasting (WRF) model. Mean historical climatology has been analyzed and contrasted with reanalysis, gridded observational datasets. Overall, VR-CESM produced comparable statistical biases to WRF in key climatological quantities. Generally, when compared with gridded observational datasets, both VR-CESM and WRF adequately represented regional climatological patterns. VR-CESM demonstrated competitive utility for studying high-resolution regional climatology when compared to a regional climate model (WRF). This assessment highlights the value of variable-resolution global climate models (VRGCMs) in capturing fine-scale atmospheric processes, projecting future regional climate and addressing the computational expense of uniform-resolution global climate models.

Supported by the satisfactory performances from the previous evaluation of the VR-CESM in regional climate modeling, VR-CESM is further applied to understand the impact of irrigation on the regional climate of California. Irrigation is an important contributor to the regional climate of heavily irrigated regions, and within the U.S. there are few regions that are as heavily irrigated as California's Central Valley. A flexible irrigation scheme with relatively realistic estimates of agricultural water use is employed and the impact of irrigation on mean historical climatology and heat extremes is investigated. It turns out that high-resolution simulations of regional climate in CESM, particularly over heavily irrigated regions, should likely enable the irrigation parameterization to better represent local temperature statistics.

Based on VR-CESM, which proves to be a useful tool for regional climate studies, we have also investigated the projected changing character of precipitation frequency and intensity in the western United States over the 21st century. Both mean changes to precipitation and distributions of both non-extreme and extreme events, projected by the VR-CESM model under climate forcing, have been investigated.  Although constrained by water influx and soil moisture, changes to extreme precipitation are hypothesized to follow the Clausius-Clapeyron relationship more closely than total precipitation amount. We aim to add value to the study of regional climate change through the use of a reliable model to reduce uncertainty in predicting changes to both mean climatology and extremes. We expect these studies can contribute to the characterization, detection and attribution of mean climatology and precipitation extremes at multi-scale.}

%{The abstract submitted as part of your dissertation, in the introductory pages, does not have a word limit. It should follow the same format as the rest of your dissertation (1.5 inch left margin, double-spaced, consecutive page numbering, etc.).}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\acknowledgments{Over all these years, there are a lot help I have received from numerous people, who have contributed significantly to the completion of this thesis. First of all, I would like to thank my advisor, Paul Ullrich, for his support and inspiration through all my Ph.D. studies. Many thanks go to my research group fellow Alan M. Rhoades for helping in modeling work and collaborating in part of my research work. I would also like to offer thanks to the many collaborators I have had the pleasure of working for the past three years, in particular, Colin M. Zarzycki at NCAR.

Specifically, for the work in chapter 2, I would like to thank Dr. Travis O'Brien for the many useful conversations on climate model assessment. We also want to thank Xue Meng Chen for her assistance in performing the WRF simulations. As for chapter 3, I'd like to thank Dr. Travis O'Brien, Dr. Richard Grotjahn and Dr. Graham E. Fogg for many useful suggestions. We would also like to thank IT support for our local UC Davis computing cluster. Lastly, for chapter 4, I would like to thank Michael Wehner for sharing the 0.25$^\circ$ uniform-resolution CESM dataset, and his many suggestions. The authors also want to thank Alan M. Rhoades for providing the simulation output and providing his feedback on the content.

Special thanks go to all the friends I have met at the University of California, Davis with whom I have experienced a lot enjoyable moments. They are Shaoyun, Marielle, Xuemeng, Jing and many others who have been my good company. Thanks for all of you showing me the joy of life. I also want to thank my best friends and the beloved ones for all the unforgettable memory.

Most importantly, I would like to thank my family for their support and encouragement all the way. Without them, I cannot get through all the difficulties I have come across all these years. Special thanks go to my father and mother, who have tried their best to provide me what I need and raised me up to where I am today.}

%Acknowledgements to those who helped you get to this point. They should be listed by chapter when appropriate.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Each chapter can be in its own file for easier editing and brought in with the \include command.
% Then use the \includeonly command to speed compilation when working on a particular chapter.
%%% \includeonly{ucdavisthesis_example_Chap1}

\begin{document}

\newcommand{\bibfont}{\singlespacing}
% need this command to keep single spacing in the bibliography when using natbib

\bibliographystyle{agufull08}
%\bibliographystyle{unsrtnat}
%many other bibliography styles are available (IEEEtran, mla, etc.). Use one appropriate for your field.

\makeintropages %Processes/produces the preliminary pages

\include{Chap1}
%\include{Chap2}
%\include{Chap3}
%\include{Chap4}
\include{Chap5}

\bibliography{ucdthesis}

%It is important to write an abstract that gives a clear description of the content and major divisions of the dissertation
% The UMI abstract uses square brackets!

\UMIabstract[Regional climate is becoming one of the most important areas under the umbrella of climate change research, as more information is now needed at finer scales. Over the past decade, variable-resolution global climate models (VRGCMs) have been introduced as an alternative way for studying regional climate and applications, contrasting to traditional Regional climate models (RCMs). In this thesis, regional climate has been studied from past to future over western United States (especially, California), working with VR-CESM--the newly developed variable-resolution enabled Community Earth System Model. In this thesis, firstly, VR-CESM is assessed for long-term regional climate modeling over California against a traditional RCM -- the Weather Research and Forcasting (WRF) model. VR-CESM demonstrated competitive utility for studying high-resolution regional climatology when compared to a regional climate model (WRF). This assessment highlights the value of variable-resolution global climate models (VRGCMs) in capturing fine-scale atmospheric processes, projecting future regional climate and addressing the computational expense of uniform-resolution global climate models. VR-CESM is further applied to understand the impact of irrigation on the regional climate of California's Central Valley, as irrigation is an important contributor to the local climate of heavily irrigated regions. A flexible irrigation scheme with relatively realistic estimates of agricultural water use is employed and the impact of irrigation on mean historical climatology and heat extremes is investigated. It turns out that high-resolution simulations of regional climate in CESM, particularly over heavily irrigated regions, should likely enable the irrigation parameterization to better represent local temperature statistics. As last main part of this thesis, the projected changing character of precipitation in the western United States over the 21st century has been investigated. Both mean changes to precipitation and distributions of both non-extreme and extreme events, projected by the VR-CESM model under climate forcing, have been studied. Changes to extreme precipitation are hypothesized and projected to follow the Clausius-Clapeyron relationship more closely than total precipitation amount. We expect these studies can contribute to the characterization, detection and attribution of mean regional climatology and precipitation extremes at multi-scale with the use of a reliable model.]

%The abstract that is submitted to UMI must be formatted as shown in the example here. {\color{red}The body of the abstract cannot exceed 350 words.} It should be in typewritten form, double-spaced, and on bond paper. {\color{red}It is important to write an abstract that gives a clear description of the content and major divisions of the dissertation}, since UMI will publish the abstract exactly as submitted. {\color{red}Students completing their requirements under Plan A should provide extra copies of the typed summary for use by the dissertation committee during the examination.}

\end{document} 